---
title: "Peer Assignment 1"
---
## Charging necessaries libraries
```{r ,results='hide',message=FALSE}
library(knitr)
library(dplyr)
library(chron)
library(reshape2)
library(lattice)
```

## Loading and preprocessing the data

Dowload file from link provided by course ([Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)) into temporal directory and unzip it in the working directory.

```{r,cache=T}
fileURL<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
temp <- tempfile()
setInternet2() #Active connections
download.file(fileURL,temp)
unzip(temp)
unlink(temp)

```

Load into a data table

```{r}
Data<-read.csv("activity.csv")
Data_NA_removed<-na.omit(Data)
```

Process data, modify table shape to get the info necessary
```{r}
reshaped_Data<-melt(Data_NA_removed,id=c("date","interval"),measure.vars="steps")
```


## What is mean total number of steps taken per day?

1. Total number of steps per day
```{r,results='asis'}
steps_per_day<-dcast(reshaped_Data,date~variable,sum)
number_of_steps<-sum(steps_per_day$steps)
```

The number of steps are `r number_of_steps`

2. Histogram of steps per day
  
```{r}
histogram<-hist(steps_per_day$steps, plot=F,breaks=30)
plot(histogram,col="darkblue",border="blue",main="Steps Histogram",xlab="Steps per day",ylab="Frec",
               ylim=NULL,xaxt='n',yaxt='n')
axis(side=2, at=seq(0,32, 2), labels=T)
axis(side=1, at=seq(0,25000, 1000), labels=T)
```

3. Mean and median of steps per day.
 
```{r}
mean_steps<-mean(steps_per_day$steps,na.rm=T)
median_steps<-median(steps_per_day$steps,na.rm=T)
```
 
The mean of steps per day is `r format(mean_steps,scientific = F)` steps.

The median of steps per day is `r median_steps` steps.

## What is the average daily activity pattern?

1.Time series plot.

```{r}
steps_per_interval<-dcast(reshaped_Data,interval~variable,sum)
plot(steps_per_interval$interval,steps_per_interval$steps,type = "l", main="Daily activity pattern", xlab=
       "5 minutes interval", ylab="Mean of number of steps",col="dark blue")
```


```{r}
interval_max_steps_number<-steps_per_interval[steps_per_interval$steps %in% 
                                                max(steps_per_interval$steps),]$interval
```

2.The interval with maximun mean of steps is the interval number `r interval_max_steps_number`.

## Imputing missing values

```{r}
NA_number<-sum(is.na(Data))
```

1.There exists `r NA_number` missing values

2.Missing values changed by mean value of the interval
```{r}
median_per_interval<-dcast(reshaped_Data,interval~variable,mean)
NA_values <- numeric()
for (i in 1:nrow(Data)) {
    obs <- Data[i, ]
    if (is.na(obs$steps)) {
        steps <- subset(median_per_interval, interval == obs$interval)$steps
    } else {
        steps <- obs$steps
    }
    NA_values<- c(NA_values, steps)
}
```

3.New dataset created with missing values filled in.
```{r}
Data_NA_filled<-Data
Data_NA_filled$steps<-NA_values
```
4. Creating an histogram.
```{r}
reshape_NA_filed_data<-melt(Data_NA_filled,id=c("date","interval"),measure.vars="steps")
steps_per_day_NA_filled<-dcast(reshape_NA_filed_data,date~variable,sum)
histogram<-hist(steps_per_day_NA_filled$steps, plot=F,breaks=30)
plot(histogram,col="darkblue",border="blue",main="Steps Histogram",xlab="Steps per day",ylab="Frec",
               ylim=NULL,xaxt='n',yaxt='n')
axis(side=2, at=seq(0,32, 2), labels=T)
axis(side=1, at=seq(0,25000, 1000), labels=T)
```


```{r}
mean_steps_NA_filled<-mean(steps_per_day_NA_filled$steps)
median_steps_NA_filled<-median(steps_per_day_NA_filled$steps)
total_steps_NA_filled<-sum(steps_per_day_NA_filled$steps)
```
 
The mean of steps per day is `r format(mean_steps_NA_filled,scientific = F)` steps.

The median of steps per day is `r format(median_steps_NA_filled,scientific=F)` steps.

There is no modification in the mean but the median has taken the value from the mean.

The total daily number of steps has raised till `r format(total_steps_NA_filled,scientific=F)` steps.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a factor varible to indicate if the date is from a weekday or a weekend.

```{r}
Data_NA_filled$Weekend=ifelse(is.weekend(Data_NA_filled$date),"weekend","weekday")
reshape_NA_filed_data_weekend<-melt(Data_NA_filled,id=c("date","interval","Weekend"),measure.vars="steps")
```

2.Make a panel plot containing a time series for weekdays and other one for weekdays.
```{r}
steps_per_interval_NA_filled<-dcast(reshape_NA_filed_data_weekend,interval+Weekend~variable,sum)
xyplot(steps ~ interval | factor(Weekend), data=steps_per_interval_NA_filled, 
       type = 'l',
       layout=c(1,2),
       main="Average Number of Steps Taken 
       \nAveraged Across All Weekday Days or Weekend Days",
       xlab="5-Minute Interval (military time)",
       ylab="Average Number of Steps Taken")
```
